Expos\'{e} Master Thesis\\ Rollup System for Layer 2 Scaling

### 2022-12-08 meeting-notes/rollup-system/main.tex
spell-checked, some cosmetics

The goal of this work is to investigate the design and implementation of a rollup system for the Ethereum blockchain. To make that palatable, we suggest a number of restrictions on the functionality:
\begin{itemize}
\item only transfers of Ether, i.e., there are no contracts on the rollup;
\item optimistic rollup (less ambitious).
\end{itemize}

@@

In rollup systems, users sign transactions and submit them to an aggregator (sequencer) that checks, orders, and executes these transactions. The aggregator then compresses the data into a rollup batch and compresses the batch with other necessary data (e.g., system state ) into a rollup block and publishes that rollup block as a single transaction on Layer 1 (L1). There are two different ways to publish the data to the network. The first approach is optimistic rollup, which assumes that the off-chain transactions are valid and does not require validity proofs for the off-chain transactions in the rollup batch. Another approach is zero-knowladge rollup (ZK rollup), in which the aggregator must publish proofs of validity. In contrast, an optimistic rollup responds to a fraud proof to determine where transactions are invalid. After a rollup batch is submitted to L1, anyone can challenge the results of a rollup transaction by computing a fraud proof.

The design should contain only the core components of a rollup system, but is capable of extending
the system if we want to enrich it for future research. For security, the system relies on a fraud
detection procedure as in optimistic rollups. A fraud proof is an assertion that a state transition
is invalid and therefore the entire batch should be reversed. The prototype targets the Etherum
blockchain. Fig~\ref{rollup-design} sketches a possible architecture of the prototype rollup system.

@@

\subsection{Interacting with the main network}
As depicted in Fig.~\ref{rollup-design}, the rollup system should include some main actors such as smart contract, aggregator, and verifier.

\subsubsection{On-chain smart contract}
The operation of the rollup system is controlled by a smart contract that is deployed on-chain. Users must deposit funds into the smart contract before they can use the system. The funds is then locked in the contract and the user receives a corresponding amount that can be used in the rollup system. The smart contract is responsible  for the following tasks.
\begin{itemize} 
\item receives and processes rollup blocks; 
\item stores and updates the status of the rollup system;  
\item processes the dispute when it receives a fraud proof from a verifier in the challenge period.
\end{itemize}

@@

\subsubsection{Verifier}
The system allows the aggregator to publish rollup batches without providing proof of validity. However, to prevent any misbehaviour by the aggregator, the rollup system proposes a period of time, called the challenge period, during which a verifier can challenge a state transition. Note that when rollup batches are published on L1, the aggregator must deposit a certain amount of funds that is locked until the end of the challenge period. Similarly, the verifier must lock the same amount as the aggregator when disputing the aggregator.

When a verifier initiates the challenge process by submitting a fraud proof, the transactions in the rollup batch are re-executed on-chain to determines who wins the challenge. If the aggregator is penalised, as a result, half of the aggregator's deposit is transferred to the verifier and the other half is burned. If the verifier fails to prove the aggregator's wrongdoing, the same will happen with the verifier's deposit.

A verifier is responsible for
\begin{itemize} 
\item watches the rollup system to detect a published batch with invalid transactions and turns it into a fraud proof;
\item starts the dispute process by submitting a fraud proof to L1 when it finds an invalid batch; 
\item cooperates with L1 in the dispute process.
\end{itemize}

\subsection{Data availability}
Since a malicious aggregator can steal funds by publishing invalid transactions, the security of the system relies on an honest verifier re-executing transactions in a rollup batch and submitting a fraud proof to challenge invalid state transitions. Since on-chain execution is based on the published data, the data availability is critical because without access to this data, the verifier cannot create fraud evidence. The data contains the transaction batch and other necessary data in a rollup block in a highly compressed form and is published on-chain as calldata for the smart contract.
%A system state is a term that refers to the available information about the rollup system at a given time. 
%The state includes key information about the network such as accounts, balances, etc.

If the full transaction data is published on the mainnet, it could be expensive. Therefore, the rollup system uses compression techniques to reduce the amount of transaction-related data. Compression of the data is achieved by publishing only the information required to deterministically re-execute the transaction.

@@

\section{Tasks}
\subsection{Survey of current rollup technologies}
\begin{itemize} 
\item Optimistic rollups. 
\item ZK rollups. 
\item The efficiency of rollups for Layer 2 scaling.
\end{itemize}
\subsection{Investment technologies}
\begin{itemize} 
\item The state tree and state root. 
\item Data compression: various techniques for compressing transaction data. 
\item Dispute progress.
\end{itemize}
\subsection{Development of rollup system}
\begin{itemize} 
\item Design, code, and deploy the smart contract on the chain. 
\item Design and implement L2 system. 
\item Develop the interface for interaction between L1 and L2.
\end{itemize}
\end{document}

>>>>

The goal of this work is to investigate the design and implementation of a rollup system for the Ethereum blockchain. To make that palatable, we suggest a number of restrictions on the functionality:
\begin{itemize}
\item only transfers of Ether, i.e., there are no contracts on the rollup;
\item optimistic rollup (less ambitious);
\item only a single aggregator (rather than multiple, to avoid the need of a consensus protocol on the roll up).
\end{itemize}

@@

In rollup systems, users sign transactions and submit them to an aggregator (sequencer) that checks, orders, and executes these transactions. The aggregator then compresses the data into a rollup batch and compresses the batch with other necessary data (e.g., system state ) into a rollup block and publishes that rollup block as a single transaction on Layer 1 (L1). There are two different ways to publish the data to the network. The first approach is optimistic rollup, which assumes that the off-chain transactions are valid and does not require validity proofs for the off-chain transactions in the rollup batch. Another approach is zero-knowledge rollup (ZK rollup), in which the aggregator must publish proofs of validity. In contrast, an optimistic rollup responds to a fraud proof to determine where transactions are invalid. After a rollup batch is submitted to L1, anyone can challenge the results of a rollup transaction by computing a fraud proof.

The design should contain only the core components of a rollup system, but it should be designed for extensibility if we want to enrich it for future research. For security, the system relies on a fraud 
detection procedure as in optimistic rollups. A fraud proof is an assertion that a state transition
is invalid and therefore the entire batch should be reversed. The prototype targets the Ethereum
blockchain. Fig~\ref{rollup-design} sketches a possible architecture of the prototype rollup system.

@@

\subsection{Interacting with the main network}
As depicted in Fig.~\ref{rollup-design}, the main actors of the rollup system are a moderating smart contract on the main chain, an aggregator, and a verifier.

\subsubsection{On-chain smart contract}
The operation of the rollup system is controlled by a smart contract that is deployed on-chain. Users must deposit funds into the smart contract before they can use the rollup. Funds are then locked in the contract and the user receives a corresponding amount of tokens on the rollup system. The smart contract is responsible  for the following tasks.
\begin{itemize} 
\item registering and deregistering users on the rollup;
\item receiving and processing rollup block summaries; 
\item storing and updating the status of the rollup system;  
\item processing disputes when a verifier submits a fraud proof in the challenge period.
\end{itemize}

@@

\subsubsection{Verifier}
The system allows the aggregator to publish rollup batches without providing proof of validity. However, to prevent any misbehaviour by the aggregator, the rollup system proposes a period of time, called the challenge period, during which a verifier can challenge a state transition. When rollup batches are published on L1, the aggregator must deposit a certain amount of funds that is locked until the end of the challenge period. Similarly, the verifier must lock the same amount as the aggregator when disputing the aggregator.

When a verifier initiates the challenge process by submitting a fraud proof, the transactions in the rollup batch are re-executed on-chain to determine who wins the challenge. If the aggregator is penalised, as a result, half of the aggregator's deposit is transferred to the verifier and the other half is burned. If the verifier fails to prove the aggregator's wrongdoing, the same will happen with the verifier's deposit.

A verifier is responsible for
\begin{itemize} 
\item watching the rollup system to detect a published batch with invalid transactions and turn it into a fraud proof;
\item starting the dispute process by submitting a fraud proof to L1 when it finds an invalid batch; 
\item cooperating with L1 in the dispute process.
\end{itemize}

\subsection{Data availability}
Since a malicious aggregator can steal funds by publishing invalid transactions, the security of the system relies on honest verifier(s) re-executing transactions in a rollup batch and submitting a fraud proof to challenge invalid state transitions. Since on-chain execution is based on the published data, the data availability is critical because without access to this data, the verifier cannot create fraud evidence. The data contains the transaction batch and other necessary data in a rollup block in a highly compressed form and is published on-chain as calldata for the smart contract.
%A system state is a term that refers to the available information about the rollup system at a given time. 
%The state includes key information about the network such as accounts, balances, etc.

Publishing the full transaction data on the mainnet could be expensive. Therefore, the rollup system uses compression techniques to reduce the amount of transaction-related data. Compression of the data is achieved by publishing only the information required to deterministically re-execute the transaction.

@@

\section{Tasks}
\subsection{Related work: Survey of current rollup technologies}
\begin{itemize} 
\item Optimistic rollups. 
\item ZK rollups. 
\item The efficiency of rollups for Layer 2 scaling.
\end{itemize}
\subsection{Investment technologies}
\begin{itemize} 
\item The state tree and state root. 
\item Data compression: various techniques for compressing transaction data. 
\item Dispute progress.
\end{itemize}
\subsection{Development of rollup system}
\begin{itemize} 
\item Design, code, and deploy the smart contract on the chain. 
\item Design and implement L2 system. 
\item Develop the interface for interaction between L1 and L2.
\end{itemize}
\end{document}
